The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-08-25 article.tex
Per recommendation: added a note that I'm selling a book about the CDK (currently not really of financial interest)
Converted a blob to a proper table

%% CTfile Format Support
%% To tabulate, r=read, w=write
%% MOLfile V2000=rw, V3000=rw
%% RXNfile V2000=rw, V3000=r
%% SDfile MOLfile V2000=rw, MOLfile V3000=r, RGfile=n/a
%% RGfile V2000=rw
%% RDfile n/a

@@

CDK \textbf{2.0} adds support for V3000 and has optimized and extended
support for V2000. Currently these are considered separate formats requiring
a user to know what version is being read beforehand. Future APIs will aim
to simplify this and provide a unified reader.

@@

\subsubsection*{Molecular fingerprints}
Also the molecular fingerprints in CDK has seen development since the 1.2
release. Traditionally CDK fingerprints were represented using the
\textit{BitSet} class shipping with Java. The benefits of that class include
such things as already implemented methods for modifying bit sets using other
bit sets. However the class keeps a vector of bits in memory. The solution was
excellent for hashed, relatively small fingerprints, \textit{e.g.}, 1024 bits,
\textit{i.e.}, 10 bits indexing space. However implementing a fingerprint
designed to avoid collisions with, \textit{e.g.} a 32 bit indexing space using
this approach would be highly memory-inefficient. To allow for multiple
fingerprint representation implementations a bit fingerprint interface was
introduced. Also, although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer vectors
supported in CDK as well. The fingerprints currently existing in CDK are listed
in Table~2\ref{tab:fingerprints}.

@@

An overview of key modules with description, important changes, and dependencies
on third-party libraries is given in Table~3\ref{tab:modules} and the dependencies
between the CDK modules are depicted in Figure~6\ref{fig:deps}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Competing interests}
JWM and NJ work for companies that sell solutions based on the CDK.

@@

  \subsection*{Table 2 - The molecular fingerprints in CDK.}
  \label{tab:fingerprints}
  Listed are the currently available molecular fingerprint in CDK with
  information about whether they come as bit / count versions, what CDK version
  they appeared in, their default size and where applicable also relevant
  references.
  \vskip 1\baselineskip

@@

      \subsection*{Table 3 - A selection of key CDK modules with major changes.}\label{tab:modules}
  An overview of a selection of often used CDK modules with description,
  dependencies on third-party libraries, and the major changes since
  version~1.2. Dependencies between modules are depicted in Figure~4\ref{fig:deps}.
  \vskip 1\baselineskip

>>>>

@@

CDK \textbf{2.0} adds support for V3000 and has optimized and extended
support for V2000. Currently these are considered separate formats requiring
a user to know what version is being read beforehand. Future APIs will aim
to simplify this and provide a unified reader. An overview of currently
supported CTfile formats is given in Table~2\ref{tab:ctfileFormats}.

@@

\subsubsection*{Molecular fingerprints}
Also the molecular fingerprints in CDK has seen development since the 1.2
release. Traditionally CDK fingerprints were represented using the
\textit{BitSet} class shipping with Java. The benefits of that class include
such things as already implemented methods for modifying bit sets using other
bit sets. However the class keeps a vector of bits in memory. The solution was
excellent for hashed, relatively small fingerprints, \textit{e.g.}, 1024 bits,
\textit{i.e.}, 10 bits indexing space. However implementing a fingerprint
designed to avoid collisions with, \textit{e.g.} a 32 bit indexing space using
this approach would be highly memory-inefficient. To allow for multiple
fingerprint representation implementations a bit fingerprint interface was
introduced. Also, although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer vectors
supported in CDK as well. The fingerprints currently existing in CDK are listed
in Table~3\ref{tab:fingerprints}.

@@

An overview of key modules with description, important changes, and dependencies
on third-party libraries is given in Table~4\ref{tab:modules} and the dependencies
between the CDK modules are depicted in Figure~6\ref{fig:deps}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Competing interests}
JWM and NJ work for companies that sell solutions based on the CDK. ELW sells
a book describing the CDK functionality.

@@

    %% CTfile Format Support
%% To tabulate, r=read, w=write
      \subsection*{Table 2 - CTfile Format Support..}\label{tab:ctfileFormats}

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lll}
  \textbf{Format}            & \textbf{V2000}  & \textbf{V3000} \\ \hline
    MOLfile & read and write & read and write \\
    RXNfile & read and write & read \\
    SDfile MOLfile & read and write & read \\ % FIXME: RGfile=n/a
    RGfile & read and write & \\
    RDfile & & \\ % FIXME: yes, we do have support
    \end{tabular}
    \end{minipage}

  \subsection*{Table 3 - The molecular fingerprints in CDK.}
  \label{tab:fingerprints}
  Listed are the currently available molecular fingerprint in CDK with
  information about whether they come as bit / count versions, what CDK version
  they appeared in, their default size and where applicable also relevant
  references.
  \vskip 1\baselineskip

@@

      \subsection*{Table 4 - A selection of key CDK modules with major changes.}\label{tab:modules}
  An overview of a selection of often used CDK modules with description,
  dependencies on third-party libraries, and the major changes since
  version~1.2. Dependencies between modules are depicted in Figure~4\ref{fig:deps}.
  \vskip 1\baselineskip
